---
title: "Assignment 9"
author: "Kumar Ramasundaram"
date: "July 10, 2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Github repository for this project is 
https://github.com/kramasundaram/Homework/blob/master/Assignment%209

## R Markdown
This file has the R code and output for assignment 9.


```{r LoadLibraries}
#Load necessary libraries for the project.
library(XML)
library(RCurl)
library(rvest)
library(ggplot2)
```

## Question 1
Scrape  the IMDB webpage http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1. Of  particular  interest  is  the  table  of  the  Cast  in  order  of  crediting.    Please  scrape  this  table and  make  it  a  data.frame() of  the  Cast  in  your  R  environment

```{r}
# Read and clean the data from IMDB website
url <- 'https://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1'
webpage <- read_html(url)

# read the cast information and store it in text format. Find the container name for the cast  using slector gadget tool. I found .itemprop span container is for the cast.
name_html <- html_nodes(webpage,'.itemprop span')
cast_data <- html_text(name_html)
head(cast_data)

# Remove any leading or trailing white spaces
cast_data <- trimws(cast_data, which = c("both"))

# Cast information in a data frame
cast_df <- as.data.frame(cast_data)
class(cast_df)
dim(cast_df)

# Read the character information. .character is the container name for character information.
charac_html <- html_nodes(webpage,'.character')
charac_data <- html_text(charac_html)
head(charac_data)

# Clean the data. First remove new line character.
charac_data <- gsub("\\n", "", charac_data)

# Remove leading and trailing white spaces
charac_data <- trimws(charac_data, which = c("both"))
charac_df <- as.data.frame(charac_data)
class(charac_df)
dim(charac_df)

# put cast and character information in a data frame
name_character <- data.frame(cast_df, charac_df)

# Question 1 c
# Remove NAs, empty rows or rows with special character
name_character <- name_character[!apply(is.na(name_character) | name_character == "" | name_character == "...", 1, all),]

# Columns names for the two columns
colnames(name_character) <- c("Actor", "Character")
head(name_character)
tail(name_character)
str(name_character)
class(name_character)
dim(name_character)

# Question 1 d
# Split  the  Actor's  name  into  twocolumns:  FirstName  andSurname.    Keep  in  mind  that  some actors/actresses  have  middle  names  as  well.    Please  make  sure  that  the  middle  names  are  in  theFirstName  column,  in  addition  to  the  first  name

# Split the actor names by space and assign only the last name to last name. Everything else is part of first name
actor_name = strsplit(as.character(name_character$Actor), ' ')
firstname = sapply(actor_name, function(x) x[1:length(x)-1])
firstname = sapply(firstname, paste, collapse = " ")
lastname = sapply(actor_name, function(x) x[length(x)])
name_df <- data.frame(firstname, lastname, name_character$Character)

# Question 1 d
# Present  the  first  10  rows  of  the data.frame()-It  should  have  only  FirstName,  Surname,  and  Character  columns.

# column names and printing first 10 records
colnames(name_df) <- c("First Name", "Last Name", "Character")
head(name_df, 10)

```
### Question 2
On  the  ESPN  website,  there  are  statistics  of  each  NBA  player.    Navigate  to  the  San  Antonio  Spurscurrent  statistics  (likely  http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs).    You  are  interested  in  the  Shooting  Statisticstable
```{r}
# Question 2 a & b
# Read Player Shooting Statistics Table
espn_table = readHTMLTable('http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs', which = 2)
head(espn_table)
tail(espn_table)
class(espn_table)

# Question 2 c
# Remove the row that has the total
espn_table <- espn_table[-c(16) ,]

# Assign Column Names
colnames(espn_table) <- c("PLAYER", "FGM", "FGA", "FG%", "3PM", "3PA" , "3P%",  "FTM",  "FTA", "FT%", "2PM",  "2PA", "2P%", "PPS", "AFG%")

# Remove the row that has column names
espn_table <- espn_table[-c(1) ,]


# Separate the player name and position
player_name = strsplit(as.character(espn_table$PLAYER), ',')
playername = sapply(player_name, function(x) x[1])
position = sapply(player_name, function(x) x[2])

# Create data frame with two new columns and other statistics from old datat frame
spurs_table <- data.frame(playername, position, espn_table[2:length(espn_table)])

# Convert the column that are numbers
col_name <- c(3:ncol(spurs_table))
spurs_table[, col_name] <- lapply(spurs_table[, col_name], function(x) as.numeric(as.character(x)))

# first 10 data
head(spurs_table)

# check the data type of the columns
str(spurs_table)

# Question 2d
player_plot <- ggplot(data = spurs_table, aes(x=spurs_table$playername, y = spurs_table$FG., fill = spurs_table$position)) + geom_bar(stat="identity", width = 0.5) + labs(title = "Player's Field Goal Percentage", x = " Player Name", y = "Field Goal Percentage", fill = "Player Position") + theme(plot.title = element_text(hjust = 0.5))
player_plot + coord_flip()

```
